---
title: "United Nations Voting Trends Analysis"
author: "Kevin Mekulu"
date: "November 2, 2017"
output: html_document
---
#ABSTRACT:
There are lots of factors that influenced voting trends at the United Nations Council let's analyze those voting trends and build simple regression models to analyze the overall voting trends which will allow to answer some hidden questions.

```{r setup, include=FALSE, warning = FALSE}
library(dplyr)
```
##1.1 Import Data Analysis
we start by importing the data set and output the data 
to get an overall idea of the type of data we are dealing
with. Then we filter important features in the data such as the name of the countries and their voting percentages(Yes or No)
```{r, include=TRUE, warning = F}
votes <- readRDS("C:/Users/jkevi_000/Downloads/votes.rds")
# Print the votes dataset
print(votes)



# Filter for votes that are "yes", "abstain", or "no"
votes %>% filter(vote <= 3)

# Add another %>% step to add a year column
votes %>%
  filter(vote <= 3) %>%
  mutate(year = 1945 + session)

# Load the countrycode package
library(countrycode)


# Convert country code 100
countrycode(100, "cown", "country.name")


# Add a country column within the mutate: votes_processed
votes_processed <- votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945, 
            country = countrycode(ccode, "cown", "country.name"))

# Print votes_processed
head(votes_processed)



# Find total and fraction of "yes" votes
votes_processed %>% summarize(total = n(),
                              percent_yes = mean(vote == 1))

# Change this code to summarize by year
votes_processed %>%
group_by(year) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1)) 

# Summarize by country: by_country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# You have the votes summarized by country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Print the by_country dataset
head(by_country)


# Sort in ascending order of percent_yes
by_country %>% arrange(percent_yes)

# Now sort in descending order
by_country %>% arrange(desc(percent_yes))

# Filter out countries with fewer than 100 votes
by_country %>%
  arrange(percent_yes) %>%
  filter(total >= 100)

# Define by_year
by_year <- votes_processed %>%
  group_by(year) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))
```
##1.2 Data Visualization
This will be the most important step of our data exploration.Let's take a look at a few plots of the voting trends of different countries, especially the percentage of "Yes" votes amongst different countries over time and see what hypothesis we can infer from those graphs.
```{r, include=TRUE, warning=FALSE}


# Load the ggplot2 package
library(ggplot2)

# Create line plot
ggplot(by_year, aes(x = year, y = percent_yes))+
  geom_line()

# Change to scatter plot and add smoothing curve
ggplot(by_year, aes(year, percent_yes)) +
  geom_point() + geom_smooth()
```
Now lets turn our attention to voting trends observed in specific countries rather than the overall voting trend. We will start by looking at the trend for the United Kingdom then pick a few other countries as well.

```{r, include=TRUE, warning=FALSE}

# Group by year and country: by_year_country
by_year_country <- votes_processed %>%
  group_by(year, country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))


# Print by_year_country
head(by_year_country)

# Create a filtered version: FR_by_year
FR_by_year <- by_year_country %>%
  filter(country == "France")

# Line plot of percent_yes over time for France only
ggplot(FR_by_year, aes(year, percent_yes)) +
  geom_line()

# Vector of four countries to examine
countries <- c("United States of America", "Cameroon",
               "France", "India")

# Filter by_year_country: filtered_4_countries
filtered_4_countries <- by_year_country %>%
  filter(country %in% countries)


# Line plot of % yes in four countries
ggplot(filtered_4_countries, aes(x = year, y = percent_yes, color = country)) +
  geom_line()

# Vector of six countries to examine
countries <- c("United States", "Cameroon",
               "France", "Japan", "Brazil", "India")

# Filtered by_year_country: filtered_6_countries
filtered_6_countries <- by_year_country %>% 
                                filter(country %in% countries)



# Vector of six countries to examine
countries <- c("United States", "Cameroon",
               "France", "Japan", "Brazil", "India")

# Filtered by_year_country: filtered_6_countries
filtered_6_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes over time faceted by country
ggplot(filtered_6_countries, aes(year, percent_yes)) +
  geom_line() +
  facet_wrap(~ country, scales = "free_y")

```

##1.3 Predicitve Modelling Using Linear Regression MOdels
We will use a linear regression model in order to examine how one variable changes with respect to another by fitting a best fit line. In this specific case we will use our regression line to describe the association between the percentage of "Yes" votes by different countries over time.

```{r, include=TRUE,warning=FALSE}

# Load the broom package
library(broom)

# Linear regression of percent_yes by year for US
US_by_year <- by_year_country %>%
  filter(country == "United States of America")
US_fit <- lm(percent_yes ~ year, US_by_year)
summary(US_fit)
```
The estimated slope of our model is -0.006 meaning that the percentage of yes votes by the United States has decreased  by a factor of 0.006 over the years. Now we could dig a little deeper to find out why but we don't have enough data to answer that question. But to verify wether or not the trend is due to chance let's look at the p-value which is 1.37*10^(-07) <<< 0.05 so clearly our model is significant.


```{r,include=TRUE,warning=FALSE}

# Create US_tidied 
US_tidied <- tidy(US_fit)
```
##1.4 Deploying Linear Regression Models for Multiple Countries
Applying linear regression to individual countries would take us forever instead, let's deploy our model for multiple countries in one go.


```{r, include=TRUE, warning=TRUE}

# Load the tidyr package
library(tidyr)

# Nest all columns besides country
nested<- by_year_country %>% group_by(country) %>% nest()


# Unnest the data column to return it to its original form
unnest(nested)

#Perform Linear Regression on each nested data frame
# Load tidyr and purrr
library(tidyr)
library(purrr)



# Perform a linear regression on each item in the data column
  nested %>% mutate(model = map(data, ~lm(percent_yes ~ year, data = .)))
  
# Load the broom package
library(broom)


# Add another mutate that applies tidy() to each model
  nested %>% mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)))%>%
  mutate(tidied = map(model,tidy))

# Add one more step that unnests the tidied column
country_coefficients <- nested %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)),
         tidied = map(model, tidy)) %>%
         unnest(tidied)

# Print the resulting country_coefficients variable
head(country_coefficients)

# Print the country_coefficients dataset
head(country_coefficients)


# Filter for only the slope terms
country_coefficients %>% filter(term == "year")
```
Now let's filter significant countries only by using our p-value strategy(p-value < 0.05)
```{r,include=TRUE,warning=FALSE}

# Filter for only the slope terms
slope_terms <- country_coefficients %>%
  filter(term == "year")

# Add p.adjusted column, then filter
slope_terms %>% mutate(p.adjusted = p.adjust(p.value) )

# Filter by adjusted p-values
filtered_countries <- country_coefficients %>%
  filter(term == "year") %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < .05)
```
Now let's look at the countries where the percentage of "Yes" votes increases the fastest.
```{r,include=TRUE,warning=FALSE}

# Sort for the countries increasing most quickly
filtered_countries %>% arrange(desc(estimate))
```
Now lets look at the countries where the percentage of "Yes" votes decreases the fastest
```{r,include=TRUE,warning=FALSE}


# Sort for the countries decreasing most quickly
filtered_countries %>% arrange(estimate)
```
#2- Conclusion
Using our simple linear regression model we were able to significantly observed voting trends during United Nations voting councils. We could further our analysis by merging new datasets with new information about specific factors that would push a country to vote yes and also make use of state-of-art machine learning algorithm for more accurate trend analysis.
